---
title: "A spatial model"
output:
  html_document:
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{css, echo=FALSE}
pre, code {white-space:pre !important; overflow-x:auto}
```

```{r general options, include = FALSE}
knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par(mgp = c(1.5, .5, 0), bty = "n", plt = c(.105, .97, .13, .97))
    else NULL
  },
  prompt = function(before, options, envir) {
    options(prompt = if (options$engine %in% c("sh", "bash")) "$ " else "> ")
  })

knitr::opts_chunk$set(margin = TRUE, message = FALSE, warning = FALSE,
                      prompt = FALSE, comment = "##", collapse = FALSE,
                      cache = FALSE, autodep = TRUE,
                      dev.args = list(pointsize = 11), fig.height = 1.5 * 3.5,
                      fig.width = 1.5 * 4.24725, fig.retina = 2, fig.align = "center")

options(width = 263)

l <- "en_US.UTF-8"
Sys.setenv(LANGAGE = l)
Sys.setlocale(locale = l)
Sys.setlocale("LC_MESSAGES", l)
```

## Utilitary functions

```{r}
discretize <- function(cdf) {
  diff(cdf / tail(cdf, 1))
}
```

Let's try it:

```{r}
discretize(pweibull(0:11, 2.24, 5.42))
```

```{r}
discretize_cond <- function(cdf) {
  rescaled <- cdf / tail(cdf, 1)
  diff(rescaled) / (1 - rescaled[-length(rescaled)])
}
```

Let's try it:

```{r}
discretize_cond(pweibull(0:11, 2.24, 5.42))
```

## A minimal SIR Model

```{r}
dmax <- 11
w <- discretize(pweibull(0:dmax, 2.24, 5.42))
```

```{r}
lambda <- function(k, w, N) {
  1 - exp(-k * w / N)
}

find_k <- function(k, w, N, R0) {
  f <- function(x) abs(N * sum(sapply(w, lambda, k = x, N = N)) - R0)
  optimize(f, k)$minimum
}

find_k(c(-5, 5), w, 9e7, 3)
```

figure:

```{r}
lwd <- 3
opar <- par(pty = "s")
col <- c("black", "#800026", "#bd0026", "#e31a1c", "#fc4e2a", "#fd8d3c", "#feb24c", "#fed976", "#ffeda0")
xs <- seq(0, 1, le = 512)
plot(xs, 1 - exp(-16 * xs), type = "l", xlim = 0:1, ylim = 0:1, asp = 1,
     xlab = "proportion of infectious in the population",
     ylab = "daily probability that a susceptible becomes infected",
     xaxs = "i", yaxs = "i", lwd = lwd, col = col[1])
abline(0, 1)
lines(xs, 1 - exp(-8 * xs), lwd = lwd, col = col[2])
lines(xs, 1 - exp(-4 * xs), lwd = lwd, col = col[3])
lines(xs, 1 - exp(-2 * xs), lwd = lwd, col = col[4])
lines(xs, 1 - exp(-1.5 * xs), lwd = lwd, col = "grey", lty = 2)
lines(xs, 1 - exp(-1.2 * xs), lwd = lwd, col = "grey", lty = 3)
lines(xs, 1 - exp(-1 * xs), lwd = lwd, col = col[5])
lines(xs, 1 - exp(-.5* xs), lwd = lwd, col = col[6])
lines(xs, 1 - exp(-.25 * xs), lwd = lwd, col = col[7])
box(bty = "o")
par(opar)
```


```{r}
tmax <- 100
R0 <- 2.5

IO <- 100
S0 <- 1e6

S <- S0
I <- matrix(0, tmax, dmax)
R <- numeric(tmax)
I[1, ] <- 1 / dmax

for (i in 1:(tmax - 1)) {
  I_tot <- w %*% I[i, ]
  N <- I_tot + S + sum(R)
  lambda <- 1 - exp(-R0 * I_tot / N)
  new_cases <- S * lambda
  S <- S - new_cases
  I[i + 1, ] <- c(new_cases, I[i, -dmax])
  R[i + 1] <- I[i, dmax]
}

R <- cumsum(R)

plot(rowSums(I))
```

## Adding ICU

IFR of symptomatics = 3%

```{r}
t_max <- 200 # duration of the simulation in days
p_hos <- .001 # probability that an infection leads to a severe case
R0 <- 2.5 
lethality <- .5 # probability to die for people in ICU
S0 <- 9.5e7
I0 <- 10
p <- .99
n <- 1e5
wbl_inf <- list(shape = 2.24, scale = 5.42) # Weibull parameters of serial int (from Nishiuria)
wbl_hos1 <- list(shape = 2, scale = 10) # Weibull parameters of time in ICU
wbl_hos <- list(shape = 1, scale = 5.5) # Weibull parameters of time in ICU
wbl_hos3 <- list(shape = 1, scale = 6.139951) # Weibull parameters of time in ICU
wbl_asp <- list(shape = 1.77, scale = 6.52) # Weibull parameters of incubation time (from Linton)
wbl_inc <- list(shape = .766, scale = 2.9) # Weibull parameters of incubation time (from Linton)
```

Serial intervals distribution, discretized:

```{r}
d_inf <- ceiling(with(wbl_inf, qweibull(p, shape, scale)))
w <- discretize(with(wbl_inf, pweibull(0:d_inf, shape, scale)))
```

Distribution of time in ICU:

```{r}
d_hos <- ceiling(with(wbl_hos, qweibull(p, shape, scale)))
u <- discretize_cond(with(wbl_hos, pweibull(0:d_hos, shape, scale)))
```

Distribution of time from contamination to hospitalization:

Monte Carlo:

```{r}
incub <- with(wbl_asp, rweibull(n, shape, scale))
hospit <- with(wbl_inc, rweibull(n, shape, scale))
iht <- incub + hospit
wbl_sev <- as.list(fitdistrplus::fitdist(iht, dweibull)$estimate)
```

```{r}
d_sev <- ceiling(with(wbl_sev, qweibull(p, shape, scale)))
h <- discretize_cond(with(wbl_sev, pweibull(0:d_sev, shape, scale)))
```

Colors:

```{r}
colors <- c(`mild infections`   = "#4daf4a",
            `severe infections` = "#377eb8",
            `ICU`               = "#e41a1c",
            `deaths`            = "#984ea3")
colors <- c(`mild infections`   = "#377eb8", # = "#fecc5c",
            `severe infections` = "#fd8d3c",
            `ICU`               = "#e31a1c",
            `deaths`            = "black")
```

Figure:

```{r fig.height = 3.5, fig.width = 4.24725}
lwd = 2
xs <- seq(0, 1.5 * max(d_inf, d_hos, d_sev), le = 512)
plot(xs, with(wbl_inf, dweibull(xs, shape, scale)), type = "l", lwd = lwd,
     col = colors["mild infections"], xaxs = "i", yaxs = "i", ylim = c(0, .18),
     xlab = "durations (days)", ylab = "density of probabilty")
lines(xs, with(wbl_hos, dweibull(xs, shape, scale)), lwd = lwd, col = colors["ICU"])
lines(xs, with(wbl_sev, dweibull(xs, shape, scale)), lwd = lwd, col = colors["severe infections"])
legend("topright", legend = c("serial intervals", "from infection to hospitalization", "ICU"),
       col = colors[! startsWith(names(colors), "d")], lwd = lwd, bty = "n")
```

```{r}
S <- numeric(t_max)
I <- matrix(0, t_max, d_inf)
J <- matrix(0, t_max, d_sev)
H <- matrix(0, t_max, d_hos)
R <- numeric(t_max)
D <- numeric(t_max)
S[1] <- S0
I[1, ] <- I0 / d_inf
```

```{r}
for (t in 1:(t_max - 1)) {
  
  I_tot <- w %*% I[t, ]
  N <- I_tot + S[t] + sum(R)
  lambda <- 1 - exp(-R0 * I_tot / N)
  lambda <- 1 - exp(-(1 * R0) * (I_tot / N))
#  Lambda <- I_tot / (S0 / R0 + I_tot)
  new_cases <- S[t] * lambda
  leave_hospital <- u %*% H[t, ]
  
  S[t + 1] <- S[t] - new_cases
  I[t + 1, ] <- c((1 - p_hos) * new_cases, I[t, -d_inf])
  J[t + 1, ] <- c(p_hos * new_cases, (1 - h[-d_sev]) * J[t, -d_sev])
  H[t + 1, ] <- c(h %*% J[t, ], (1 - u[-d_hos]) * H[t, -d_hos])
  R[t + 1] <- I[t, d_inf] + (1 - lethality) * leave_hospital
  D[t + 1] <- lethality * leave_hospital
}
```

```{r}
I <- rowSums(I)
#I <- rowSums(t(t(I) * w))
J <- rowSums(J)
H <- rowSums(H)
R <- cumsum(R)
D <- cumsum(D)
```

```{r fig.height = 1 * 3.5,fig.width = 1 * 4.24725}
lwd <- 2
plot(I, lwd = lwd, col = colors["mild infections"], type = "l", ylim = c(0, 1e5),
     xlab = "time (day)", ylab = "number of people")
lines(J, lwd = lwd, col = colors["severe infections"])
lines(H, lwd = lwd, col = colors["ICU"])
lines(D, lwd = lwd, col = colors["deaths"])
legend("topright", legend = names(colors), col = colors, bty = "n", lwd = 2)
```

```{r}
tail(R, 1)
tail(D, 1)
```

## Discriminating ICU from other hospitals

```{r}
t_max <- 200 # duration of the simulation in days
p_hos <- .001 # probability that an infection leads to a severe case
R0 <- 2.5 
lethality <- .5 # probability to die for people in ICU
S0 <- 9.5e7
I0 <- 10
p <- .99
n <- 1e5
wbl_inf <- list(shape = 2.24, scale = 5.42) # Weibull parameters of serial int (from Nishiuria)
wbl_hos1 <- list(shape = 2, scale = 10) # Weibull parameters of time in ICU
wbl_hos <- list(shape = 1, scale = 5.5) # Weibull parameters of time in ICU
wbl_hos3 <- list(shape = 1, scale = 6.139951) # Weibull parameters of time in ICU
wbl_asp <- list(shape = 1.77, scale = 6.52) # Weibull parameters of incubation time (from Linton)
wbl_inc <- list(shape = .766, scale = 2.9) # Weibull parameters of incubation time (from Linton)
```

Serial intervals distribution, discretized:

```{r}
d_inf <- ceiling(with(wbl_inf, qweibull(p, shape, scale)))
w <- discretize(with(wbl_inf, pweibull(0:d_inf, shape, scale)))
```

Distribution of time in ICU:

```{r}
d_hos <- ceiling(with(wbl_hos, qweibull(p, shape, scale)))
u <- discretize_cond(with(wbl_hos, pweibull(0:d_hos, shape, scale)))
```

Distribution of time from contamination to hospitalization. Monte Carlo:

```{r}
incub <- with(wbl_asp, rweibull(n, shape, scale))
hospit <- with(wbl_inc, rweibull(n, shape, scale))
iht <- incub + hospit
wbl_sev <- as.list(fitdistrplus::fitdist(iht, dweibull)$estimate)
```

The distribution:

```{r}
d_sev <- ceiling(with(wbl_sev, qweibull(p, shape, scale)))
h <- discretize_cond(with(wbl_sev, pweibull(0:d_sev, shape, scale)))
```

```{r}
S <- numeric(t_max)
I <- matrix(0, t_max, d_inf)
J <- matrix(0, t_max, d_sev)
H <- matrix(0, t_max, d_hos)
R <- numeric(t_max)
D <- numeric(t_max)
S[1] <- S0
I[1, ] <- I0 / d_inf
```

```{r}
for (t in 1:(t_max - 1)) {
  
  I_tot <- w %*% I[t, ]
  N <- I_tot + S[t] + sum(R)
  lambda <- 1 - exp(-R0 * I_tot / N)
  lambda <- 1 - exp(-(1 * R0) * (I_tot / N))
#  Lambda <- I_tot / (S0 / R0 + I_tot)
  new_cases <- S[t] * lambda
  leave_hospital <- u %*% H[t, ]
  
  S[t + 1] <- S[t] - new_cases
  I[t + 1, ] <- c((1 - p_hos) * new_cases, I[t, -d_inf])
  J[t + 1, ] <- c(p_hos * new_cases, (1 - h[-d_sev]) * J[t, -d_sev])
  H[t + 1, ] <- c(h %*% J[t, ], (1 - u[-d_hos]) * H[t, -d_hos])
  R[t + 1] <- I[t, d_inf] + (1 - lethality) * leave_hospital
  D[t + 1] <- lethality * leave_hospital
}
```

```{r}
I <- rowSums(I)
#I <- rowSums(t(t(I) * w))
J <- rowSums(J)
H <- rowSums(H)
R <- cumsum(R)
D <- cumsum(D)
```

```{r fig.height = 1 * 3.5,fig.width = 1 * 4.24725}
lwd <- 2
plot(I, lwd = lwd, col = colors["mild infections"], type = "l", ylim = c(0, 1e5),
     xlab = "time (day)", ylab = "number of people")
lines(J, lwd = lwd, col = colors["severe infections"])
lines(H, lwd = lwd, col = colors["ICU"])
lines(D, lwd = lwd, col = colors["deaths"])
legend("topright", legend = names(colors), col = colors, bty = "n", lwd = 2)
```

```{r}
tail(R, 1)
tail(D, 1)
```
